---
title: "ETC3550/5550 Applied forecasting"
---

```{r}
#| label: load_packages
#| include: false
#| message: false
#| warning: false
#| echo: false
library(tidyverse)
options(knitr.kable.NA = '')
```

## Lecturer/Chief Examiner

* [**Rob J Hyndman**](https://robjhyndman.com). Email: [Rob.Hyndman@monash.edu](mailto:Rob.Hyndman@monash.edu)

## Tutors

* [**Mitchell O'Hara-Wild**](https://mitchelloharawild.com). Email: [Mitch.O'Hara-Wild@monash.edu](mailto:Mitch.O'Hara-Wild@monash.edu)
* Elena Sanina
* Xiaoqian Wang
* Yangzhuoran (Fin) Yang
* Zhixiang (Elvis) Yang

## Consultations

* Rob
* Mitch
* Elena
* Elvis
* Fin
* Xiaoqian

## Weekly schedule

* Pre-recorded lectures: 1 hour per week ([Slides](https://github.com/robjhyndman/fpp3_slides))
* In-person lectures: 9am Fridays, [Central 1 Lecture Theatre, 25 Exhibition Walk](https://maps.app.goo.gl/RKdmJq2tBfw8ViNT9).
* Tutorials: 1.5 hours in class per week


```{r}
#| label: schedule2
#| message: false
#| warning: false
#| echo: false
#| output: asis
# Week of mid-semester break
mid_semester_break <- 6
# Set up schedule of topics
schedule <- tibble(
  Week = seq(12),
  Date = seq(as.Date("2024-02-26"), by = "1 week", length.out = 12),
  Topic = c(
    "Introduction to forecasting and R",
    "Time series graphics",
    "Time series decomposition",
    "The forecaster's toolbox",
    "Exponential smoothing",
    "Exponential smoothing",
    "ARIMA models",
    "ARIMA models",
    "ARIMA models",
    "Multiple regression and forecasting",
    "Dynamic regression",
    "Dynamic regression"
  ),
  Chapter = c(
    "1. Getting started",
    "2. Time series graphics",
    "3. Time series decomposition",
    "5. The forecaster's toolbox",
    "8. Exponential smoothing",
    "8. Exponential smoothing",
    "9. ARIMA models",
    "9. ARIMA models",
    "9. ARIMA models",
    "7. Time series regression models",
    "10. Dynamic regression models",
    "10. Dynamic regression models"
  ),
  Chapter_URL = c(
    "https://OTexts.com/fpp3/intro.html",
    "https://otexts.com/fpp3/graphics.html",
    "https://otexts.com/fpp3/decomposition.html",
    "https://otexts.com/fpp3/toolbox.html",
    "https://otexts.com/fpp3/expsmooth.html",
    "https://otexts.com/fpp3/expsmooth.html",
    "https://otexts.com/fpp3/arima.html",
    "https://otexts.com/fpp3/arima.html",
    "https://otexts.com/fpp3/arima.html",
    "https://otexts.com/fpp3/regression.html",
    "https://otexts.com/fpp3/dynamic.html",
    "https://otexts.com/fpp3/dynamic.html"
  ),
  Assessments = c(
    "Assignment 1",
    "",
    "Assignment 2",
    "",
    "",
    "Assignment 3",
    "",
    "",
    "Assignment 4",
    "",
    "",
    "Project"
  ),
  Assessment_URL = c(
    "https://learning.monash.edu/mod/quiz/view.php?id=1343906",
    "",
    "https://learning.monash.edu/mod/assign/view.php?id=1343917",
    "",
    "",
    "https://learning.monash.edu/mod/assign/view.php?id=1343934",
    "",
    "",
    "https://learning.monash.edu/mod/assign/view.php?id=1343951",
    "",
    "",
    "https://learning.monash.edu/mod/assign/view.php?id=1343999"
  )
) |>
  mutate(
    Date = if_else(Week < mid_semester_break, Date, Date + 7)
  )
# Add in mid-semester break
bind_rows(
  schedule |> filter(Week < mid_semester_break),
  tibble(
    Week = NA,
    Date = min(schedule$Date) + 7*(mid_semester_break-1),
    Topic = "**Mid-semester break**",
    Chapter = NA,
    Chapter_URL = NA,
    Assessments = NA,
    Assessment_URL = NA
  ),
  schedule |> filter(Week >= mid_semester_break)
) |>
  mutate(
    Date = format(Date, "%d %b"),
    Topic = if_else(is.na(Week), Topic, glue::glue("[{Topic}](./week{Week}.html)")),
    Chapter = if_else(is.na(Week), Chapter, glue::glue("[{Chapter}]({Chapter_URL})")),
    Assessments = if_else(is.na(Week), Assessments, glue::glue("[{Assessments}]({Assessment_URL})"))
  ) |>
  select(-Chapter_URL, -Assessment_URL) |>
  knitr::kable(format = "markdown")
```

## Assessments

Final exam 60%, project 20%, other assignments 20%

## R package installation

Here is the code to install the R packages we will be using in this unit.

```r
install.packages(c("tidyverse","fpp3", "GGally"), dependencies = TRUE)
```
