---
title: "ETC3550/5550 Applied forecasting"
---

```{r}
#| label: load_packages
#| include: false
#| message: false
#| warning: false
#| echo: false
#| cache: false
library(tidyverse)
options(knitr.kable.NA = '')
source(here::here("course_info.R"))
week <- as.integer(1 + ((Sys.Date() - as_date(start_semester))/7))
```

## Lecturer/Chief Examiner

* [**Rob J Hyndman**](https://robjhyndman.com). Email: [Rob.Hyndman@monash.edu](mailto:Rob.Hyndman@monash.edu)

## Tutors

* [**Mitchell O'Hara-Wild**](https://mitchelloharawild.com). Email: [Mitch.O'Hara-Wild@monash.edu](mailto:Mitch.O'Hara-Wild@monash.edu)
* Elena Sanina
* Xiaoqian Wang
* Zhixiang (Elvis) Yang

## Consultations

* Rob
* Mitch
* Elena
* Elvis
* Xiaoqian

## Weekly schedule

* Pre-recorded videos: approximately 1 hour per week [[Slides](https://github.com/robjhyndman/fpp3_slides)]
* Tutorials: 1.5 hours in class per week
* Seminars: 9am Fridays, [Central 1 Lecture Theatre, 25 Exhibition Walk](https://maps.app.goo.gl/RKdmJq2tBfw8ViNT9).


```{r}
#| label: schedule2
#| message: false
#| warning: false
#| echo: false
#| output: asis
schedule |>
  transmute(
    Date = format(Date, "%d %b"),
    Show = !is.na(Week) & (Week <= week | Week <= 1),
    Topic = if_else(!Show, Topic, glue::glue("[{Topic}](./week{Week}/index.html)")),
    Chapter = if_else(is.na(Week), Chapter, glue::glue("[{Chapter}]({Chapter_URL})")),
    Assessments = if_else(is.na(Assignment), Assignment, glue::glue("[{Assignment}]({File})"))
  ) |>
  select(-Show) |>
  knitr::kable(format = "markdown")
```

## Assessments

Final exam 60%, project 20%, other assignments 20%

## R package installation

Here is the code to install the R packages we will be using in this unit.

```r
install.packages(c("tidyverse","fpp3", "GGally"), dependencies = TRUE)
```
