---
title: "Week 5: Exponential smoothing"
---

```{r}
week <- 5
source("course_info.R")
```

## What you will learn this week

* Simple exponential smoothing
* Corresponding ETS models

## Pre-class activities

Read [Sections 8.1-8.4 of the textbook](https://otexts.com/fpp3/expsmooth.html) and watch all embedded videos

## Exercises (on your own or in tutorial)

Complete Exercises 1-6, 8-9, 11-12 from [Section 5.11 of the book](https://otexts.com/fpp3/toolbox-exercises.html).

```{r}
#| output: asis
show_slides(week)
```

## Seminar activities

Try forecasting the Chinese GDP from the `global_economy` data set using an ETS model.

Experiment with the various options in the `ETS()` function to see how much the forecasts change with damped trend, or with a Box-Cox transformation. Try to develop an intuition of what each is doing to the forecasts.

[Hint: use `h=20` when forecasting, so you can clearly see the differences between the various options when plotting the forecasts.]



```{r}
#| output: asis
show_assignments()
```
