---
title: "Week 2: Time series graphics"
---

```{r}
#| cache: false
week <- 2
source("course_info.R")
```

## What you will learn this week

* Different types of plots for time series including time plots, season plots, subseries plots, lag plots and ACF plots
* The difference between seasonal patterns and cyclic patterns in time series
* What is "white noise" and how to identify it.

## Pre-class activities

Read [Chapter 2 of the textbook](https://otexts.com/fpp3/graphics.html) and watch all embedded videos

## Exercises (on your own or in tutorial)

Complete Exercises 1-5 from [Section 2.10 of the book](https://otexts.com/fpp3/graphics-exercises.html).


```{r}
#| output: asis
show_slides(week)
```

## Seminar activities

1. We have introduced various functions for time series graphics include `autoplot()`, `gg_season()`, `gg_subseries()`, `gg_lag()` and `ACF`. Use these functions to explore the quarterly tourism data for the Snowy Mountains.

   ```r
   snowy <- tourism |> filter(Region == "Snowy Mountains")
   ```
    
   What do you learn?

2. Which time plot corresponds to which ACF plot?

```{r}
#| label: acf-quiz
#| fig.asp: 0.5
#| fig.width: 12
#| echo: false
#| warning: false
#| out.width: 100%
#| cache: false
source("slides/setup.R")
library(patchwork)
cowtemp <- as_tsibble(fma::cowtemp)
USAccDeaths <- as_tsibble(USAccDeaths)
AirPassengers <- as_tsibble(AirPassengers)
mink <- as_tsibble(fma::mink)
tp1 <- autoplot(cowtemp, value) +
  labs(x = "", y = "chirps per minute", title = "1. Daily temperature of cow")
tp2 <- autoplot(USAccDeaths, value) +
  labs(x = "", y = "thousands", title = "2. Monthly accidental deaths")
tp3 <- autoplot(AirPassengers, value) +
  labs(x = "", y = "thousands", title = "3. Monthly air passengers")
tp4 <- autoplot(mink, value) +
  labs(x = "", y = "thousands", title = "4. Annual mink trappings")
acfb <- ACF(cowtemp, value) |> autoplot() +
  labs(x = "", title = "B") + ylim(-0.4, 1)
acfa <- ACF(USAccDeaths, value) |> autoplot() +
  labs(x = "", title = "A") + ylim(-0.4, 1)
acfd <- ACF(AirPassengers, value) |> autoplot() +
  labs(x = "", title = "D") + ylim(-0.4, 1)
acfc <- ACF(mink, value) |> autoplot() +
  labs(x = "", title = "C") + ylim(-0.4, 1)
(tp1 | tp2 | tp3 | tp4) / (acfa | acfb | acfc | acfd)
```

```{r}
#| output: asis
#| echo: false
show_assignments()
```
