---
title: "Week 1: What is forecasting?"
---

```{r}
week <- 1
source("course_info.R")
```

## What you will learn this week

* How to think about forecasting from a statistical perspective
* What makes something easy or hard to forecast?
* Using the `tsibble` package in R

## Pre-class activities

* Install R and RStudio on your personal computer. Instructions are provided at [https://otexts.com/fpp3/appendix-using-r.html](https://otexts.com/fpp3/appendix-using-r.html).
* Read [Chapter 1 of the textbook](http://OTexts.com/fpp3/intro.html) and watch all embedded videos

## What is forecasting?

<iframe width="100%" height="415" src="https://www.youtube.com/embed/HNJYRf0mvxg?si=k0wfI3Sq68TPm4Ek" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture; web-share" allowfullscreen></iframe>

```{r}
#| output: asis
show_slides(week)
```

## In-class activities

Your task this week is to make sure you are familiar with R, RStudio and the tidyverse packages. If you've already done ETC1010, then you may not need to do anything! But if you're new to R and the tidyverse, then you will need to get yourself up-to-speed.

Work through the first five chapters of the **LearnR** tutorial at [learnr.numbat.space](https://learnr.numbat.space). Do as much of it as you think you need. For those students new to R, it is strongly recommended that you do all five chapters. For those who have previously used R, concentrate on the parts where you feel you are weakest.


```{r}
#| output: asis
show_assignments()
```
