---
title: "Week 12: Dynamic regression"
---

```{r}
week <- 12
source(here::here("course_info.R"))
```

## What you will learn this week

* How to combine regression models with ARIMA models to form dynamic regression models
* Dynamic harmonic regression to handle complex seasonality
* Lagged predictors

## Pre-class activities

Read [Chapter 10 of the textbook](https://otexts.com/fpp3/dynamic.html) and watch all embedded videos

## Exercises (on your own or in tutorial)

Complete Exercises 2-6 from [Section 10.7 of the book](https://otexts.com/fpp3/dynamic-exercises.html).

[**Solutions to Exercises**](ex12-sol.qmd)

```{r}
#| output: asis
show_slides(week)
```

[Code used in lecture](seminar_code.R)

# Tuesday revision class

<iframe src='https://docs.google.com/gview?url=https://af.numbat.space/week12/revision_slides.pdf&embedded=true' width='100%' height=465></iframe>
<a href="https://af.numbat.space/week12/revision_slides.pdf" class='badge badge-small badge-red'>Download slides</a>

## Post-class activities

* Do any exercises not yet finished.
* Complete [past exams](/exams).
* Re-read [the textbook](https://otexts.com/fpp3/).
* Listen again to all lectures, workshops, tutorials.


```{r}
#| output: asis
show_assignments(week)
```
