---
title: "Week 2: Time series graphics"
---

```{r}
week <- 2
source(here::here("course_info.R"))
```

## What you will learn this week

* Different types of plots for time series including time plots, season plots, subseries plots, lag plots and ACF plots
* The difference between seasonal patterns and cyclic patterns in time series
* What is "white noise" and how to identify it.

## Pre-class activities

Read [Chapter 2 of the textbook](https://otexts.com/fpp3/graphics.html) and watch all embedded videos

## Monday lecture

```{r}
#| output: asis
show_slides(week)
```

<!-- * [Code used in lecture](lecture_code.R) -->

## Tuesday workshop

```{r}
#| output: asis
#show_activity(week)
show_quiz(week)
```

<!-- * [Code used in workshop](workshop_code.R) -->

## Exercises (on your own or in tutorial)

* Complete Exercises 1-11 from [Section 2.10 of the book](https://otexts.com/fpp3/graphics-exercises.html).
<!-- * [Solutions to Exercises](ex2-sol.qmd) -->

```{r}
#| output: asis
show_assignments(week)
```
