---
title: ETC3550/ETC5550 Applied forecasting
author: "Week 2: Time series graphics"
format:
  presentation-beamer:
    knitr:
      opts_chunk:
        dev: "cairo_pdf"
    pdf-engine: pdflatex
    fig-width: 7.5
    fig-height: 3.5
    include-in-header: ../header.tex
    keep_tex: yes
---

```{r setup, include=FALSE}
source(here::here("setup.R"))
library(patchwork)

global_economy <- global_economy |>
  select(Year, Country, GDP, Imports, Exports, Population)
tourism <- tourism |>
  mutate(
    State = recode(State,
      "Australian Capital Territory" = "ACT",
      "New South Wales" = "NSW",
      "Northern Territory" = "NT",
      "Queensland" = "QLD",
      "South Australia" = "SA",
      "Tasmania" = "TAS",
      "Victoria" = "VIC",
      "Western Australia" = "WA"
    )
  )
```


## Seasonal or cyclic?

\alert{Differences between seasonal and cyclic patterns:}

  * seasonal pattern constant length; cyclic pattern variable length
  * average length of cycle longer than length of seasonal pattern
  * magnitude of cycle more variable than magnitude of seasonal pattern

\pause

\begin{alertblock}{}
The timing of peaks and troughs is predictable with seasonal data, but unpredictable in the long term with cyclic data.
\end{alertblock}

## Your turn

We have introduced various functions for time series graphics include `autoplot()`, `gg_season()`, `gg_subseries()`, `gg_lag()` and `ACF`. Use these functions to explore the quarterly tourism data for the Snowy Mountains.

```r
snowy <- tourism |> filter(Region == "Snowy Mountains")
```

What do you learn?


## Trend and seasonality in ACF plots

- When data have a trend, the autocorrelations for small lags tend to be large and positive.
- When data are seasonal, the autocorrelations will be larger at the seasonal lags (i.e., at multiples of the seasonal frequency)
- When data are trended and seasonal, you see a combination of these effects.


## Which is which?

```{r, fig.height=6, fig.width=12, echo=FALSE, warning=FALSE, out.width="14.5cm"}
cowtemp <- as_tsibble(fma::cowtemp)
USAccDeaths <- as_tsibble(USAccDeaths)
AirPassengers <- as_tsibble(AirPassengers)
mink <- as_tsibble(fma::mink)
tp1 <- autoplot(cowtemp, value) +
  labs(x = "", y = "chirps per minute", title = "1. Daily temperature of cow")
tp2 <- autoplot(USAccDeaths, value) +
  labs(x = "", y = "thousands", title = "2. Monthly accidental deaths")
tp3 <- autoplot(AirPassengers, value) +
  labs(x = "", y = "thousands", title = "3. Monthly air passengers")
tp4 <- autoplot(mink, value) +
  labs(x = "", y = "thousands", title = "4. Annual mink trappings")
acfb <- ACF(cowtemp, value) |> autoplot() +
  labs(x = "", title = "B") + ylim(-0.4, 1)
acfa <- ACF(USAccDeaths, value) |> autoplot() +
  labs(x = "", title = "A") + ylim(-0.4, 1)
acfd <- ACF(AirPassengers, value) |> autoplot() +
  labs(x = "", title = "D") + ylim(-0.4, 1)
acfc <- ACF(mink, value) |> autoplot() +
  labs(x = "", title = "C") + ylim(-0.4, 1)
(tp1 | tp2 | tp3 | tp4) / (acfa | acfb | acfc | acfd)
```
