---
title: "Week 4: Simple forecasting methods"
---

```{r}
week <- 4
source(here::here("course_info.R"))
```

## What you will learn this week

* Forecasting workflow
* Four benchmark forecasting methods that we will use for comparison
* Linear trends, dummy seasonality
* Forecasting with transformations
* Forecasting with decompositions

## Pre-class activities

Read [Chapter 5 of the textbook](https://otexts.com/fpp3/toolbox.html), Sections 5.1 -- 5.3, 5.6 -- 5.7, and watch all embedded videos

Read [Chapter 7.4 of the textbook](https://otexts.com/fpp3/useful-predictors.html), and watch the embedded video

## Exercises (on your own or in tutorial)

Complete Exercises 1-5, 11 from [Section 5.11 of the book](https://otexts.com/fpp3/toolbox-exercises.html).

Complete Exercises 2 and 6 from [Section 7.10 of the book](https://otexts.com/fpp3/regression-exercises.html)

<!-- [**Solutions to Exercises**](ex4-sol.qmd) -->

```{r}
#| output: asis
show_slides(week)
```

[Code used in lecture](lecture4.R)

```{r}
#| output: asis
show_activity(week, show_solutions = FALSE)
show_assignments(week)
```
