---
title: "Week 4: The forecaster's toolbox"
---

```{r}
week <- 4
source(here::here("course_info.R"))
```

## What you will learn this week

* Four benchmark forecasting methods that we will use for comparison
* Fitted values, residuals
* Forecasting with transformations

## Pre-class activities

Read [Chapter 5 of the textbook](https://otexts.com/fpp3/toolbox.html) and watch all embedded videos

## Exercises (on your own or in tutorial)

Complete Exercises 1-5, 9-10 from [Section 3.7 of the book](https://otexts.com/fpp3/decomposition-exercises.html).

```{r}
#| output: asis
show_slides(week)
```

## Seminar activities

 * Produce forecasts using an appropriate benchmark method for household wealth (`hh_budget`). Plot the results using `autoplot()`.
 * Produce forecasts using an appropriate benchmark method for Australian takeaway food turnover (`aus_retail`). Plot the results using `autoplot()`.

  * Compute seasonal naïve forecasts for quarterly Australian beer production from 1992.
  * Test if the residuals are white noise. What do you conclude?

 * Create a training set for household wealth (`hh_budget`) by withholding the last four years as a test set.
 * Fit all the appropriate benchmark methods to the training set and forecast the periods covered by the test set.
 * Compute the accuracy of your forecasts. Which method does best?
 * Repeat the exercise using the Australian takeaway food turnover data (`aus_retail`) with a test set of four years.

```{r}
#| output: asis
show_assignments(week)
```
