---
title: "Activities: Week 3"
editor: source
engine: knitr
filters:
  - live
  - quiz-teachr
ojs-engine: true
webr:
  packages:
    - qlcheckr
    - fpp3
  autoload-packages: false
  repos:
    - https://repo.r-wasm.org/
    - https://learnr-academy.github.io/qlcheckr
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}

# Was there a COVID baby boom?

```{r}
#| include: false
vic_births <- fpp3::aus_births |> 
  filter(State == "VIC") |>
  select(-State)
```

The following data shows the number of births in Victoria each month from January 1975 to December 2021.

```{webr}
#| exercise: births
vic_births
```

```{webr}
#| exercise: births
vic_births |> autoplot(Births)
vic_births |> gg_season(Births)
```

The see-saw seasonal pattern is due to month length variation.

So let's remove it by taking the average of the daily births over the month.
First we create a tibble with the number of days in each month, then we join it to the `vic_births` data so we can compute the daily average for each month.

```{webr}
months <- tibble(
  month_name = month.abb,
  month_length = c(31, 28, 31, 30, 31, 30, 31, 31, 30, 31, 30, 31)
)
months
vic_births <- vic_births |>
  mutate(month_name = month(Month, label = TRUE)) |>
  left_join(months, by = "month_name") |>
  mutate(daily_average = Births / month_length) 
vic_births
```


```{r}
vic_births |> autoplot(daily_average)
vic_births |> gg_season(daily_average)
```

* Has the peak month for having babies changed compared to the previous `gg_season` plot? Why?
* Why has the unusual fluctuation after January 2020 apparently increased in size compared to the previous `gg_season` plot?

Let's do an STL decomposition. Experiment with the value of `window` and `robust` to see how the seasonal component changes over time. We need `robust  = TRUE` here so the unusual behaviour near the end of the series does not have a strong effect on the trend or seasonal components

```{r}
dmp <- vic_births |>
  model(stl = STL(daily_average ~ season(window = 19), robust = TRUE))
components(dmp) |> autoplot()
```

```{r}
components(dmp) |> gg_season(season_year)
components(dmp) |> gg_subseries(season_year)
```

How has the seasonal pattern changed over time?

```{r}
components(dmp) |>
  filter(Month >= yearmonth("2020 Jan")) |>
  autoplot(remainder)
```

Once we remove the 

# STL decomposition: Canadian gas data

This exercise uses the `canadian_gas` data (monthly Canadian gas production in billions of cubic metres, January 1960 – February 2005).

* Plot the data using `autoplot()`, `gg_subseries()` and `gg_season()` to look at the effect of the changing seasonality over time.
* Do an STL decomposition of the data. You will need to choose a seasonal window to allow for the changing shape of the seasonal component.
* How does the seasonal shape change over time? 
* Can you produce a plausible seasonally adjusted series?

# STL decomposition: Canberra public transport usage data

* Exam 2024 B2.
