---
title: "Activities: Week 8"
execute:
  echo: true
  output: false
  eval: false
---


## AUD exchange rates

The data set contains Australian dollar exchange rates from the start of 2023 to 22 April 2025.

```{r}
#| label: load-aud-xr
library(fpp3)
aud_xr <- readRDS(url("https://bit.ly/aud_xr"))
```

1. How has the tsibble been indexed? Why wouldn't `Date` be used for the index?

2. See what currencies are included:

   ```{r}
   #| label: currencies
   aud_xr |>
     distinct(Currency, Description)
   ```

3. Choose a currency and plot its exchange rate using `gg_tsdisplay` like this. Does it look stationary?

   ```{r}
   aud_xr |>
     filter(Currency == ???) |>
     autoplot(Rate)
   ```

4. Now plot the first difference of the exchange rate. Does it look stationary?

   ```{r}
   aud_xr |>
     filter(Currency == ???) |>
     autoplot(difference(Rate))
   ```

5. Repeat for another couple of exchange rates. Does differencing always result in a stationary series? Why?

6. If we used the `unitroot_ndiffs()` function, would it give the same answer?

   ```{r}
   aud_xr |>
     features(Rate, unitroot_ndiffs)
   ```

## Tasmanian accommodation takings

Let's extract the Tasmanian data from `aus_accommodation`. Check the help to see what this data set contains.

```{r}
tas_takings <- aus_accommodation |>
  filter(State == "Tasmania") |>
  select(Takings)
```

We will try to find transformations that make this series stationary.

1. First, try to stabilise the variance using a Box-Cox transformation.

   ```{r}
   tas_takings |>
     autoplot(
       box_cox(Takings, lambda = ???)
     )
   ```

2. Then, try to stabilise the mean using seasonal differencing.

   ```{r}
   tas_takings |>
     autoplot(
       box_cox(Takings, lambda = ???) |>
       difference(lag = ???)
     )
   ```

3. If that doesn't work, try a first difference as well.

   ```{r}
   tas_takings |>
     autoplot(
       box_cox(Takings, lambda = ???) |>
       difference(lag = ???) |>
       difference(lag = ???)
     )
   ```

   The final transformed series should appear stationary.

4. Check it's ACF.

   ```{r}
   tas_takings |>
     ACF(
       box_cox(Takings, lambda = ???) |>
       difference(lag = ???) |>
       difference(lag = ???)
     ) |>
     autoplot()
   ```

   What do you conclude?

5. Now use the `unitroot_ndiffs` and `unitroot_nsdiffs` functions to check what differences they would suggest:

   ```{r}
   tas_takings |>
     features(
       box_cox(Takings, lambda = ???),
       unitroot_nsdiffs
     )
   ```

   ```{r}
   tas_takings |>
     features(
       box_cox(Takings, lambda = ???) |> difference(lag = ???),
       unitroot_nsdiffs
     )
   ```

   Does it suggest the same number of differencing as you did? Why/why not?
